%\documentclass[conference]{IEEEtran}
\documentclass[twocolumn,twoside]{IEEEtran}

\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
% \usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{tabularx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{listings}

\usepackage[T1]{fontenc}
\usepackage{times}

%\usepackage{biblatex} %Imports biblatex package
%\addbibresource{bibliography.bib} %Import the bibliography file

\usepackage[numbers]{natbib}

\graphicspath{{images/}}


\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\begin{document}

\lstdefinestyle{mystyle}{
    basicstyle=\ttfamily\footnotesize,
    breakatwhitespace=false,    
    xleftmargin=4pt,     
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2,
    keywordstyle=\color{black}\bfseries
}

\lstset{style=mystyle}



\title{
Recording the Lifecycle of Software Models
}

\author{\IEEEauthorblockN{Brendan P. Bonner}\\
\IEEEauthorblockA{\textit{School of Computing} \\
\textit{Dublin City University}\\
Dublin, Ireland \\
brendan.bonner2@mail.dcu.ie}

}

\maketitle

\begin{abstract}
Artificial Intelligence is built on trustworthiness of the systems that provide the intelligence. The foundation of this is built upon our understanding of the role of software models in this process. Similar to measuring social trust in humans, we can only increase our levels of trust by being exposed to the full competence and vulnerabilities of the environment to get to the point in time where trust can be assessed. The pathway for the human brain to grow and evolve is recorded in genealogy, education and experience. In this paper, we investigate the feasibility and benefits for replicating this process for machine learning systems. During this process, we developed an ecosystem where AI systems can have their entire lifecycle recorded and made available for scrutiny. The framework for a verifiable chain of provenance of the lifecycle of artificial intelligence development may be a first step into addressing shortcomings AI research has which potentially could stymie innovation in certain critical systems that need public support. The outcome of the research cannot enforce a system to be trustworthy, instead providing additional insight into a software models \textit{personality}, without needed to directly examine each neuron. 
\end{abstract}

\begin{IEEEkeywords}
Artificial Intelligence, Provenance, Explainability, Model Analysis, Trust, Cloud\end{IEEEkeywords}

\input{1.Introduction}

\input{2.Background}

\input{3.Method}

\input{4.Results}

\input{5.Conclusions}

\bibliographystyle{IEEEtran}
\bibliography{bibliography}


\end{document}

